---
title: Summary of Maxnet modeling and conservation gap analysis results per taxon
author: Colin Khoury, Daniel Carver
output:
  html_document:
  code_folding: hide
highlight: tango
theme: yeti
toc: no
toc_depth: 4
toc_float:
  collapsed: yes
smooth_scroll: yes
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# function for flitering list based on character values
include <- function (theList, toMatch){
  matches <- unique (grep(paste(toMatch,collapse="|"),
                          theList, value=TRUE))
  return(matches)
}
```

## Summary of the models for `r species`.

The table below shows a summary of the counts for the important categories in the data. A useful record is one that has a complete set of latitude and longitude coordinates associated with it. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# taxa is the

baseDir <- paste0(sp_dir)
csv <- list.files(baseDir, pattern = ".csv", recursive = TRUE , full.names = TRUE)
if (file.exists(paste0(baseDir,'/counts.csv'))){
  counts <- as.data.frame(read.csv(paste0(baseDir,'/counts.csv'))) %>%
    dplyr::select(c("totalRecords", "totalUseful", "totalGRecords", "totalGUseful","totalHRecords", "totalHUseful"))
  colnames(counts) <- c("Total Records", "Total Useful", "Total G Records", "Total G Useful","Total H Records", "Total H Useful")
}else{
  print("There are no samples present for this species")
}
DT::datatable(counts)
```
<br>
<br>
<br>
  
### Predictors Used in Modeling Process 
The table below shows the variables listed in order of importance. As correlated variables were removed, the “Included in the Modeling Process” column defines if the variable was used in the final modeling process.  
```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"predictorImportance.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  t2 <- read.csv(include(csv,"predictorImportance.csv"))
  t2 <- dplyr::left_join(x = t2, y= layerDescription, by = c("varNames" = "layer"))
  t2 <- t2 %>% dplyr::select("varNames","description" ,"importance","includeInFinal","source")
    colnames(t2) <- c("Variable Name", "Description", "Relative Importance to Model", "Included in the Modeling Process", "Data Source")
  DT::datatable(t2)
}
```

### Evaluation metrics
To account for random variation occurring during modeling, the process was run ten times (K=10) for all species with more than 25 points. Species with more than 10 but less than 25 occurrences were run six times, and those with < 10 were modeled three times. The table below shows the statistical results for each run.

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("No models were produced for this species due to insufficent number of samples")
}else{
  # evalMetric <- include(csvs,"eval_metrics_rep.csv")
  modeledData <- read.csv(include(csv, "presBackgroundWithBiovars.csv"))
  totalVals <- nrow(modeledData)
  eval <- read.csv(include(csv,"eval_metrics_rep.csv"))
  kfold <- nrow(eval)
  if(kfold == 10){
    training <- round(totalVals * 0.9, 0)
    testing <- round(totalVals * 0.1, 0)
  }
  if(kfold == 6){
    training <- round(totalVals * 0.83, 0)
    testing <- round(totalVals * 0.17, 0)
  }
  if(kfold == 3){
    training <- round(totalVals * 0.67, 0)
    testing <- round(totalVals * 0.33, 0)
  }
  
  
  
  
  eval$Species <- species
  eval$Testing <- testing
  eval$Training <- training
  eval$`Background + Training` <- testing + training
  eval <- eval %>% 
    dplyr::select("Species" , "Training", "Testing","Background + Training", "AUCtrain", "sensi_train","speci_train","threshold_train", "max.TSS_train","minROCdist_train","method_train", "AUCtest","threshold_test","sensi_test"       , "speci_test","matthews.cor_test", "LR_pos_test"      
, "LR_neg_test","kappa_index_test" )
  colnames(eval) <- c("Species" , "Training", "Testing", "Background + Training", "AUC Train", "Sensitivity Train","Specificity Train","Threshold Train", "Maximun TSS Train","Minimun ROC Distance Train","Threshold Method Train", "AUC Test","Threshold Test","Sensitivity Test"       
, "Specificity Test","Matthews Correlation Test", "LR_pos_test"      
, "LR_neg_test","kappa_index_test" )
  DT::datatable(eval)
}
```
<br>
<br>
<br>

  
### Statistics for Median of the Model Runs
The median result across the replicates is shown below. Per previous articles we have done, to be considered an accurate and stable model, we are looking for an AUC >= 0.7; SDAUC < 0.15; and ASD15 <= 10

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  # evalMetric <- include(csvs,"eval_metrics_rep.csv")
  eval2 <- read.csv(include(csv,"eval_metrics.csv"))
  eval2 <- eval2 %>%
    dplyr::select(species, ATAUC, STAUC, ASD15, VALID)
  
  eval2$Testing <- testing
  eval2$Training <- training
  eval2$`Background + Training` <- testing + training
  eval2 <- eval2 %>%
    dplyr::select(species,Training,Testing, `Background + Training`,ATAUC, STAUC, ASD15, VALID)
  colnames(eval2) <- c("Species","Training", "Testing", "Background + Training", "AUC","SDAUC","ASD15", "Valid")
  DT::datatable(eval2)
}
```
<br>
<br>
<br>
  
```{r echo=FALSE, message=FALSE, warning=FALSE}

if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  tif <- list.files(path = baseDir, pattern = '.tif', recursive = TRUE, full.names = TRUE)
      median <- raster(include(tif, "prj_median"))
      thrshold <- raster(include(tif, "spdist_thrsld_median"))
      sd <- raster(include(tif, "prj_std"))
      mess <- raster(include(tif, "messMapThres"))
      kDense <- try(raster(include(tif, "kernalDensity")))
    }
tmap_mode("view")
tmap_options(basemaps = "OpenStreetMap")
```

### Maps of model outputs.


#### Median

The map below shows the median spatial result from the maxent model runs. Values in the key refer to the probability of occurrence. The occurrences used in the modeling process are represented by black dots and the attribute data associated with the points can be viewed by clicking on the dots.
```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  tm_shape(median) + tm_raster()+
    tm_shape(cleanPoints) + tm_dots()+
    tm_scale_bar() 
}
```

<br>

  
#### Standard Deviation
Map of the standard deviation result across maxent model runs. Areas with a high standard deviation imply a limited agreement between the runs. This is most likely to occur in areas with low sample (occurrence point) density.
```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  qtm(shp = sd)
}
```

<br>
  
  
#### Threshold
Map of the threshold (binary presence-absence) result from the maxent models. This is the map used to create the genus level predicted species richness map.


```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  palette1 <- c(  "#FFFFFF","#45B320")
  tm_shape(thrshold) + tm_raster(palette=palette1)+
    tm_scale_bar() 
}
```

#### Mess Map
The mess map identifies regions of the model that over project the top predictor. That is, areas of predicted presences that are above or below the observed range of the values assocaited with known occurrence data. There is less confidence that these areas are part of the distribution of the species because they are beyond the know conditions at what the species has been observed.
Areas with value 0 equal species potential range.
Areas with value 1 are regions that are potential over projected.
Areas with value 2 are predicted presences locations within the know environmental limit of the species. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  mess[is.na(mess)] <- 0
  mess1 <- mess + thrshold
  palette1 <- c(  "#FFFFFF","#8c19b5", "#45B320")
  tm_shape(mess1) + tm_raster(palette=palette1)+
    tm_scale_bar() 
  
  
}
```

#### Density of Sampling locations 
The map below show the relative concentration of occurence data across the predicted suitable habitat for the species. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(tif, "kernalDensity")) == 0){
  print("No kernal density raster was generated for this species")
}else{
  th2 <- thrshold
  th2[th2 == 0] <- NA
  kd1 <- kDense * th2
  tm_shape(kd1) + tm_raster()+
    tm_scale_bar() 
}
```

### Conservation gap analysis

#### Ex situ Conservation

The table below shows the ex situ conservation summary. SRSex is a gross comparison of germplasm (G) and reference (H) records. GRSex analyzes how comprehensively the G records cover the maxent model spatially. ERSex analyzes how well the G records cover the maxent model with regard to ecosystems covered. 
All of the conservation metrics are on a scale from 0-100, with 0 = poor conservation and 100 = perfectly sufficient conservation. The final ex situ conservation score is called FCSex and is a mean of the three ex situ conservation scores.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# exSummary <- include(csv, "exsitu/summary.csv")
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("There are no G points present so this process could not be completed.")
}else{
tableEx <- dplyr::bind_cols(counts, read.csv(include(csv, "exsitu/summary.csv")))
tableEx <- tableEx[c(7,1,2,3,4,5,6,8,9,10, 11)]
names(tableEx) <- c("ID","Total Records","Total Useful","Total G Records", 
                    "Total G Useful","Total H Records","Total H Useful","SRSex", "GRSex","ERSex","FCSex")
DT::datatable(tableEx)
}
```
<br>
<br>
<br>


```{r echo=FALSE, message=FALSE, warning=FALSE}
if(counts$`Total G Useful` == 0){
  print("There are no G points present so this process could not be completed.")
}else{
  # pull in all g points.
  #No idea why this is not working, going back to old indexing
  # %>%   filter(type == "G")
  occList <- cleanPoints[which(cleanPoints$type == "G"),]
  if(nrow(occList) != 0){
    sp = sp::SpatialPoints(occList[,c("longitude", "latitude")])
   gapRaster <- raster(include(tif, "ga50"))
   gapRaster2 <- raster(include(tif, "grs_pa_PAs_narea_areakm2"))
   gapRaster2a <- gapRaster2[is.na(gapRaster2)] <- 0
   collgap1 <- thrshold + gapRaster
   GRSin <- thrshold + gapRaster2
  }else{
print("There are no G points present so this process could not be completed.")
  } 
}
    
```

#### GRSex
Map of the potential distribution, with previous germplasm collection points surrounded by a 50 km buffer overlaid. Only germplasm points are displayed on the map.

Areas of Native Range = 0

Potential Distribution = 1

Area Where samples have been collected = 2

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("There are no G points present so this process could not be completed.")
}else{
  
  occList <- cleanPoints[which(cleanPoints$type == "G"),]
  if(nrow(occList) == 0){
  print("There are no G points present so this process could not be completed.")
  }else{ gPoint <- cleanPoints[cleanPoints@data$type == "G",]
  palette1 <- c(  "#FFFFFF","#45B320","#7570b3")
  if(length(gPoint@data)>0){
    thes1 <- thrshold 
    thes1[thes1==0]<-NA 
    bufferInThres <- collgap1 * thes1
    bufferInThres[is.na(bufferInThres)]<-0
    comb2 <-  bufferInThres + thrshold
    gGaps <- thrshold - bufferInThres 
    writeRaster(x = gGaps, filename = paste0(sp_dir,'/gap_analysis/exsitu/gapMap.tif'),overwrite=TRUE)
  tm_shape(comb2) + tm_raster(palette = palette1)+
  tm_shape(gPoint) + tm_dots()+
    tm_scale_bar()
  }else{
palette2 <- c(  "#FFFFFF","#45B320")
tm_shape(collgap1) + tm_raster(palette=palette2)+
    tm_scale_bar()
    }
  }
}

```

<br>
<br>
<br>



#### In Situ Conservation


The table below shows the in situ conservation summary.  SRSin reports the proportion of occurrences which fall within protected areas (WDPA database 2019). GRSin analyzes how comprehensively protected areas cover the maxent model spatially. ERSin analyzes how well protected areas cover the maxent model with regard to ecosystems covered. 
All of the conservation metrics are on a scale from 0-100, with 0 = poor conservation and 100 = perfectly sufficient conservation. The final in situ conservation score is called FCSin and is a mean of the three in situ conservation scores.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# inSummary <- include(csv, "insitu/summary.csv")
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not able to be ran on this species due to insufficent number of samples")
}else{
tableIn <- dplyr::bind_cols(counts, read.csv(include(csv, "insitu/summary.csv")))
tableIn <- tableIn[c(7,1,2,3,4,5,6,11,12,13,14)]
names(tableIn) <- c("ID","Total Records","Total Useful","Total G Records",
                    "Total G Useful","Total H Records","Total H Useful","SRSin",
                    "GRSin","ERSin","FCSin")
DT::datatable(tableIn)
}
```
<br>
<br>
<br>

#### GRSin

Map of the potential distribution, with distribution occurring within existing protected areas (WDPA 2019) highlighted.

Areas of Native Range = 0

Potential Distribution = 1

Protected Lands within the Potential Distribution = 2

```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not able to be ran on this species due to insufficent number of samples")
}else{
gapRaster2 <- proArea %>%
  raster::crop(thrshold) %>%
  raster::mask(thrshold)

      if(file.exists(include(tif, "spdist_thrsld_median"))){
        gapRaster2[is.na(gapRaster2)] <- 0 
        thres1 <- thrshold
        thres1[thres1==0]<-NA
        gapRaster2a <- gapRaster2 * thres1
        gapRaster2a[is.na(gapRaster2a)]<-0
        comb2a <- gapRaster2a + thrshold 
        palette3 <- c(  "#FFFFFF","#45B320","#7570b3")
tm_shape(comb2a) + tm_raster(palette = palette3)
      }else{
        print("Models were not able to be ran on this species due to insufficent number of samples")
      }
}

```
<br>
<br>
<br>

#### Combined Summary 
This table shows the combined ex situ and in situ conservation metrics. FCSc-mean is the final combined conservation score. We categorize taxa based on the final combined scores (0-25 = high priority (HP) for further conservation work; 25-50 medium (MP); 50-75 low (LP); and 75-100 sufficiently conserved (SC)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# inSummary <- include(csv, "insitu/summary.csv")
if(length(include(csv,"eval_metrics_rep.csv")) == 0){
  print("Models were not are to be ran on this species due to insufficent number of samples")
}else{
  tableCo<- dplyr::bind_cols(counts, read.csv(include(csv, "combined/fcs_combined.csv")))
  tableCo <- tableCo[c(7,1,2,3,4,5,6, 8,9,10,11,12,13,14,15)]
  colnames(tableCo) <- c("Species", "Total Records", "Total Useful", "Total G Records", "Total G Useful", "Total H Records", "Total H Useful", "FCSex","FCSin"          
, "FCSc_min","FCSc_max","FCSc_mean"      
,"FCSc_min_class" , "FCSc_max_class","FCSc_mean_class")
DT::datatable(tableCo)
}
```
<br>
<br>
<br>

This document was last last updated on `r Sys.Date()`

