---
title: Summary of maxent modeling results and conservation gap analysis results per taxon
author: Colin Khoury, Daniel Carver
output:
  html_document:
  code_folding: hide
highlight: tango
theme: yeti
toc: no
toc_depth: 4
toc_float:
  collapsed: yes
smooth_scroll: no
---
  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# function for flitering list based on character values
include <- function (theList, toMatch){
  matches <- unique (grep(paste(toMatch,collapse="|"),
                          theList, value=TRUE))
  return(matches)
}
```


## Species Richness Map
This map combines all threshold maps for individual species.


```{r  echo=FALSE, message=FALSE, warning=FALSE}
n= 1
speciesList <- species
vector <- c("feature")
baseDir <- "D:/cwrOfNA/gap_analysis/Vaccinium"
### need to put is a clause incase
for(i in speciesList){
  tif <- list.files(path = paste0(baseDir,"/",i,"/",run_version),
                    pattern = '.tif', recursive = TRUE, full.names = TRUE)
  thres <- include(tif, "spdist_thrsld")
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}

removeNA = function(rasterPath){
  r1 <- raster::raster(rasterPath)
  r1[is.na(r1)] <- 0
  return(r1)
}


rasterList <- lapply(X = vector, FUN = removeNA)


extend_all =function(rasters){
  extent(Reduce(extend,rasters))
}

sum_all =
  function(rasters, extent){
    re = lapply(rasters, function(r){extend(r, extent, value=0)})
    Reduce("+",re)
  }
r_sum = sum_all(rasterList, extend_all(rasterList))
#Replace all zeros with NA
r_sum[r_sum == 0]<-NA
qtm(r_sum)
raster::writeRaster(x = r_sum, filename = paste0(outputDir , '/speciesRichness'), format = "GTiff",overwrite=TRUE)

```
<br>
<br>

  
## Summary of Conservation Statistics for all Species
  
The SRS conservation statistic is based on all known occurance records for the species (Total_records). The other conservation metrics are based on the expert evaluated dataset that was used in the modeling process (Records_Used_in_Model).   


```{r  echo=FALSE, message=FALSE, warning=FALSE}

modelingData <- read.csv("path to location of modeling data")

species_summary <- function(species){
  
  count_dir <- paste0(base_dr,"/",species)
  gap_dir <- paste0(base_dr,"/",species,"/",runFolder,"/gap_analysis")
  
  numberUsedInModel <- modelingData %>%
    filter(Taxon_final == species)%>%
    nrow()
  
  colsNames <- c( "Scientific_Name", "Total_records",
                  "Total_with_coords", "Total_G", "Total_G_with_coords", "Total_H", "Total_H_with_coords",
                  "Total_records_Tunisia","Total_with_coords_Tunisia", "Total_G_Tunisia",
                  "Total_G_with_coords_Tunisia","Total_H_Tunisia", "Total_H_with_coords_Tunisia",
                  "SRS", "GRS_ex", "ERS_ex", "FCS_ex", "GRS_in", "ERS_in", "FCS_in", "FCSc_min",
                  "FCSc_max", "FCSc_mean", "Priority_category")
  
  summ_table <- as.data.frame(matrix(ncol = length(colsNames), nrow = 1))
  colnames(summ_table) <- colsNames
  
  ### if country specific add theses
  # "Total_records_Tunisia","Total_with_coords_Tunisia", "Total_G_Tunisia",
  # "Total_G_with_coords_Tunisia","Total_H_Tunisia", "Total_H_with_coords_Tunisia",
  
  summ_table$Scientific_Name <- as.character(species)
  
  if(file.exists(paste0(count_dir, "/counts.csv"))){
    counts <- read.csv(paste0(count_dir, "/counts.csv"), header = TRUE)
    summ_table$Total_records <- counts$totalRecords
    summ_table$Total_with_coords <- counts$totalUseful
    summ_table$Total_G <- counts$totalGRecords
    summ_table$Total_G_with_coords <- counts$totalGUseful
    summ_table$Total_H <- counts$totalHRecords
    summ_table$Total_H_with_coords <- counts$totalHUseful
    #summ_table$Records_Used_in_Model <- numberUsedInModel
  }else{
    summ_table$Total_records <- NA
    summ_table$Total_with_coords <- NA
    summ_table$Total_G <- NA
    summ_table$Total_G_with_coords <- NA
    summ_table$Total_H <- NA
    summ_table$Total_H_with_coords <- NA
    #summ_table$Records_Used_in_Model <- NA
  }
  if(file.exists(paste0(count_dir, "/countsTunOnly.csv"))){
    counts <- read.csv(paste0(count_dir, "/countsTunOnly.csv"), header = TRUE)
    summ_table$Total_records_Tunisia <- counts$totalRecords
    summ_table$Total_with_coords_Tunisia <- counts$totalUseful
    summ_table$Total_G_Tunisia <- counts$totalGRecords
    summ_table$Total_G_with_coords_Tunisia <- counts$totalGUseful
    summ_table$Total_H_Tunisia <- counts$totalHRecords
    summ_table$Total_H_with_coords_Tunisia <- counts$totalHUseful
    
  }else{
    summ_table$Total_records_Tunisia <- NA
    summ_table$Total_with_coords_Tunisia <- NA
    summ_table$Total_G_Tunisia <- NA
    summ_table$Total_G_with_coords_Tunisia <- NA
    summ_table$Total_H_Tunisia <- NA
    summ_table$Total_H_with_coords_Tunisia <- NA
    
  }
  # if (is.na(pri2$VALID[which(pri2$Species==species)])){
  #   summ_table$Model_type <- "NA"
  # } else if(pri2$VALID[which(pri2$Species==species)]){
  #   summ_table$Model_type <- "SDM"
  # }else if (pri2$VALID[which(pri2$Species==species)]==FALSE){
  #   summ_table$Model_type <- "CA50"
  # }
  
  
  if(file.exists(paste0(gap_dir,"/exsitu/summary.csv"))){
    exsitu <- read.csv(paste0(gap_dir,"/exsitu/summary.csv"), header = TRUE)
    summ_table$SRS <- round(exsitu$SRS, 2)
    summ_table$GRS_ex <- round(exsitu$GRS, 2)
    summ_table$ERS_ex <- round(exsitu$ERS, 2)
    summ_table$FCS_ex <- round(exsitu$FCS, 2)
  }
  
  if(file.exists(paste0(gap_dir,"/insitu/summary.csv"))){
    insitu <- read.csv(paste0(gap_dir,"/insitu/summary.csv"), header = TRUE)
    summ_table$GRS_in <- round(insitu$GRS, 2)
    summ_table$ERS_in <- round(insitu$ERS, 2)
    summ_table$FCS_in <- round(insitu$FCS, 2)
  }
  
  
  
  if(file.exists(paste0(gap_dir,"/combined/fcs_combined.csv"))){
    combined <- read.csv(paste0(gap_dir,"/combined/fcs_combined.csv"), header = TRUE)
    summ_table$FCSc_min <- round(combined$FCSc_min, 2)
    summ_table$FCSc_max <- round(combined$FCSc_max, 2)
    summ_table$FCSc_mean <- round(combined$FCSc_mean, 2)
    summ_table$Priority_category <- as.character(combined$FCSc_mean_class)
  }
  
  return(summ_table)
  
}

species_table<-lapply(1:length(speciesList), function(i){
  pt<-species_summary(speciesList[i])
  return(pt)
  
})

species_table <- do.call(rbind, species_table)
utils::write.csv(species_table, paste0(outputDir, "/species_summary_",Sys.Date() ,".csv"), row.names = FALSE, quote = FALSE)
DT::datatable(species_table, options = list(pageLength = 25))

```



<br>
  <br>
  <br>
  
  
  ## Median Model Runs
  Evaluation statistics for the median model run for species that enough occurence points to be modeled.

```{r  echo=FALSE, message=FALSE, warning=FALSE}
n= 1
speciesList <- species
vector <- c("feature")
### need to put is a clause incase
for(i in speciesList){
  csv <- list.files(path = paste0(base_dr,"/",i,"/",runFolder,"/"), pattern = '.csv', recursive = TRUE, full.names = TRUE)
  thres <- include(csv, "eval_metrics.csv")
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}
df <- lapply(vector, read.csv)%>%
  bind_rows
df <-df[-c(8),]
write.csv(df, paste0(outputDir, "/median_summary_",Sys.Date() ,".csv"), row.names = FALSE, quote = FALSE)
DT::datatable(df, options = list(pageLength = 25))

```

<br>
  <br>
  <br>
  
  <!-- ## Priority locations for further collecting -->
  <!-- This richness map shows the overlapping potential distributions of taxa - only where they have not yet been collected previously (i.e. not within 50 km of an existing germplasm accession point). High values are areas where many species might be found and have not yet been collected, and therefore may represent geographic targets for efficient collecting in the future -->
  
  
  ```{r  echo=FALSE, message=FALSE, warning=FALSE}
#define mask

# n= 1
#   ### need to put is a clause incase
# for(i in speciesList){
# tif <- list.files(path = paste0(base_dr,"/",i,"/",runFolder,"/"), pattern = '.tif', recursive = TRUE, full.names = TRUE)
# thres <- include(tif, "spdist_thrsld")
# g50 <- include(tif, "ca50_g_narea_pa")
# mask <- include(tif, "narea_mask")
#
# # gapRaster2 <- raster(include(tif, "grs_pa_PAs_narea_areakm2"))
# # GRSin <- thres + gapRaster2
#
# if(length(thres) > 0){
#   thres1 <- raster(thres) - raster(g50)
#   thres1 <- crop(thres1, raster(mask))
#   thres1[is.na(thres1)] <- 0
#   if(n==1){
#     vector <- thres1
#   }else{
# vector <- vector + thres1
# }
# n = n+1
# }
# }
#
# tmap_mode("view")
# #Replace all zeros with NA
# vector[vector == 0]<-NA
# qtm(vector)
# raster::writeRaster(x = vector, filename = paste0(outputDir , '/exsituGapRichness'), format = "GTiff")


```
