---
title: Summary of maxent modeling results and conservation gap analysis results per run
author: Colin Khoury, Daniel Carver
output:
  html_document:
  code_folding: hide
highlight: tango
theme: yeti
toc: no
toc_depth: 4
toc_float:
  collapsed: yes
smooth_scroll: yes
---

this will change but based on inital content for CSSA this is what I want to summarise. 
 - richness map 250 or all 
 - number of species in each category md-
 - gap maps 
 - counts for EOO AOO 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# function for flitering list based on character values
include <- function (theList, toMatch){
  matches <- unique (grep(paste(toMatch,collapse="|"),
                          theList, value=TRUE))
  return(matches)
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# define run 
run_version  <- 'test20191023'
print(paste0("This is a summary of the ",run_version," modeling outputs."))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# write out a list of non modeling species 
# lessThen3Points <- lowOccurence[!is.na(lowOccurence)]
# lessThen10Points <- notModeled[!is.na(notModeled)]
# 
# write.csv(x = lessThen3Points, file = paste0(base_dir, "/runSummaries/lessThen3Points.csv"))
# write.csv(x = lessThen10Points, file = paste0(base_dir, "/runSummaries/lessThen10Points.csv"))
```
### Species Richness map 
 We were able to model 364 species from 65 genera 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Pull all genuslevel summary rasters 
n= 1
vector <- c()

### need to put is a clause incase
for(i in genera){
  tif <- list.files(path = paste0(gap_dir,"/",i),
                    pattern = '.tif', recursive = TRUE, full.names = TRUE)
  thres <- include(tif, "richnessMap")
  thres <- thres[length(thres)] # richness map is output by date. This is select the last element in the list which is the most recent. 
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}

# set extent equal to NAshp 
rasterList <- lapply(X = vector, FUN = removeNA)

# all all rasters together 
r_sum = sum_all(rasterList, extend_all(rasterList))
#Replace all zeros with NA
r_sum[r_sum == 0]<-NA

# save content 
try(qtm(r_sum))
try(writeRaster(x = r_sum, filename = paste0(base_dir,"/runSummaries/richnessMap.tif"), overwrite=TRUE))


```


### Counts per gap ananlysis categories

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Pull all species level fcsSummary docs 
files <- list.files(path = gap_dir , pattern = 'fcs_combined.csv', recursive = TRUE, full.names = TRUE)
files <- include(files, run_version)

# create empty dataframe with species, Difference between EX and IN, and summary priority 
df <- data.frame()
# select columns for each species and rbind to df 
for(i in 1:length(files)){
  file <- read.csv(files[i]) %>%
    dplyr::select(ID, FCSex, FCSin, FCSc_mean_class)
  df <- rbind(df,file)
}
# save content 

# generate summaries of 
# number of species in each group (lp,Mp,hp) and number which ex or in is higher. 
counts <- df %>% 
  mutate(diff = FCSex -FCSin)
for(i in 1:nrow(counts)){
  if(counts$diff[i] > 0){
    counts$ExsituOverInsitu[i] <- 1
  }else{
        counts$ExsituOverInsitu[i] <- 0
  }
}
write.csv(x = counts, file = paste0(base_dir, "/runSummaries/gapStatistics.csv"))

c2 <- counts %>%
  group_by(FCSc_mean_class, ExsituOverInsitu) %>%
  dplyr::summarise(count = n())
DT::datatable(c2)
write.csv(x = c2, file = paste0(base_dir, "/runSummaries/gapStatisticsSummaries.csv"))

```

### Gap Maps 

Insitu gap map 
- took the species richness map generated about and masked out all protected areas. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Insitu 

# Pull in protect areas map and convert all 1 to NA 
p1 <- proArea %>%
  raster::crop(naSHP) %>%
  raster::mask(naSHP)
p1[is.na(p1)] <- 10
p1[p1==1] <- NA 
p1[p1==10] <- 1

# mUTLIPLe NA protected areas by species richness map 
inGap <- p1 * r_sum 

try(qtm(inGap))
try(writeRaster(x = inGap, filename = paste0(base_dir,"/runSummaries/insituGap.tif"), overwrite=TRUE))
```

### Exsitu gap map 
so this one is a little goofy. Of the 360ish species we were able to model only 121 species had any useful g points. 
- I expect there is an error within the data transformation step that is causing this. For example only 1 of cucurbita species had any g points. Not the case with our more quality dataset. 
- I don't suggest spending much time on this map. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# exsitu 
files <- list.files(path = gap_dir , pattern = 'gapMap.tif', recursive = TRUE, full.names = TRUE)
files <- include(files, run_version)

# compline all these rasters 
# set extent equal to NAshp 
rasterList1 <- lapply(X = files, FUN = removeNA)

# all all rasters together 
g_sum = sum_all(rasterList1, extend_all(rasterList1))
#Replace all zeros with NA
g_sum[g_sum == 0]<-NA

# save content 
try(qtm(g_sum))
try(writeRaster(x = g_sum, filename = paste0(base_dir,"/runSummaries/gapMap.tif"), overwrite=TRUE))

# set extent equal to NAshp 

# add all of them together. 

```



### EOO AOO summary 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Pull all species level listingValues csvs 
files <- list.files(path = gap_dir , pattern = 'listingValues.csv', recursive = TRUE, full.names = TRUE)
files <- include(files, run_version)

# compile into a single dataframe 
df <- data.frame()
for(i in 1:length(files)){
  df <- rbind(df,read.csv(files[i]))
}
write.csv(x = df, file = paste0(base_dir, "/runSummaries/eooAooData.csv"))
# summarise for counts of each. 
Eoo <- df %>%
  group_by(EOO.Status, AOO.Status) %>%
  dplyr::summarise('Number of Species' = n())

sumVal <- sum(Eoo$`Number of Species`)
for(i in 1:nrow(Eoo)){
  Eoo$`Percentage of Species`[i] <- (Eoo$`Number of Species`[i]/sumVal)*100
}


DT::datatable(Eoo,options = list(pageLength = 15) )
write.csv(x = Eoo, file = paste0(base_dir, "/runSummaries/eooAooDataSummary.csv"))
```