---
title: Summary of maxent modeling results and conservation gap analysis results per taxon
author: Colin Khoury, Daniel Carver
output:
  html_document:
  code_folding: hide
highlight: tango
theme: yeti
toc: no
toc_depth: 4
toc_float:
  collapsed: yes
smooth_scroll: no
---
  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# function for flitering list based on character values
include <- function (theList, toMatch){
  matches <- unique (grep(paste(toMatch,collapse="|"),
                          theList, value=TRUE))
  return(matches)
}

# function for replacing na values with 0 
removeNA = function(rasterPath){
  r1 <- raster::raster(rasterPath)
  r1[is.na(r1)] <- 0
  return(r1)
}

# function for extending all rasters to an equal extent
extend_all =function(rasters){
  extent(Reduce(extend,rasters))
}

# function for adding all rasters together
sum_all = function(rasters, extent){
    re = lapply(rasters, function(r){extend(r, extent, value=0)})
    Reduce("+",re)
  }


```


## Species Richness Map
This map combines all threshold maps for individual species.


```{r  echo=FALSE, message=FALSE, warning=FALSE}

n= 1
vector <- c()
baseDir <- paste0(gap_dir, "/", genus)
### need to put is a clause incase
for(i in speciesList){
  tif <- list.files(path = paste0(baseDir,"/",i,"/",run_version),
                    pattern = '.tif', recursive = TRUE, full.names = TRUE)
  thres <- include(tif, "spdist_thrsld")
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}


rasterList <- lapply(X = vector, FUN = removeNA)




r_sum = sum_all(rasterList, extend_all(rasterList))
#Replace all zeros with NA
r_sum[r_sum == 0]<-NA
try(qtm(r_sum), FALSE)
try(raster::writeRaster(x = r_sum, filename = paste0(baseDir, "/richnessMap_",Sys.Date() ,".tif"), overwrite=TRUE))

```
<br>
<br>

  
## Summary of Conservation Statistics for all Species
  
The SRS conservation statistic is based on all known occurance records for the species (Total_records). The other conservation metrics are based on the expert evaluated dataset that was used in the modeling process (Records_Used_in_Model).   


```{r  echo=FALSE, message=FALSE, warning=FALSE}

modelingData <-occData

species_summary <- function(species){
  
  count_dir <- paste0(baseDir,"/",species,"/",run_version)
  gapDir <- paste0(baseDir,"/",species,"/",run_version, "/gap_analysis")
  
  numberUsedInModel <- modelingData %>%
    filter(taxon == species)%>%
    nrow()
  
  colsNames <- c( "Scientific_Name", "Total_records",
                  "Total_with_coords", "Total_G", "Total_G_with_coords", "Total_H", "Total_H_with_coords",
                  "SRS", "GRS_ex", "ERS_ex", "FCS_ex", "GRS_in", "ERS_in", "FCS_in", "FCSc_min",
                  "FCSc_max", "FCSc_mean", "Priority_category")
  
  summ_table <- as.data.frame(matrix(ncol = length(colsNames), nrow = 1))
  colnames(summ_table) <- colsNames
  
  summ_table$Scientific_Name <- as.character(species)
  
  if(file.exists(paste0(count_dir, "/counts.csv"))){
    counts <- read.csv(paste0(count_dir, "/counts.csv"), header = TRUE)
    summ_table$Total_records <- counts$totalRecords
    summ_table$Total_with_coords <- counts$totalUseful
    summ_table$Total_G <- counts$totalGRecords
    summ_table$Total_G_with_coords <- counts$totalGUseful
    summ_table$Total_H <- counts$totalHRecords
    summ_table$Total_H_with_coords <- counts$totalHUseful
    #summ_table$Records_Used_in_Model <- numberUsedInModel
  }else{
    summ_table$Total_records <- NA
    summ_table$Total_with_coords <- NA
    summ_table$Total_G <- NA
    summ_table$Total_G_with_coords <- NA
    summ_table$Total_H <- NA
    summ_table$Total_H_with_coords <- NA
    #summ_table$Records_Used_in_Model <- NA
  }
  
  if(file.exists(paste0(gapDir,"/exsitu/summary.csv"))){
    exsitu <- read.csv(paste0(gapDir,"/exsitu/summary.csv"), header = TRUE)
    summ_table$SRS <- round(exsitu$SRS, 2)
    summ_table$GRS_ex <- round(exsitu$GRS, 2)
    summ_table$ERS_ex <- round(exsitu$ERS, 2)
    summ_table$FCS_ex <- round(exsitu$FCS, 2)
  }
  
  if(file.exists(paste0(gapDir,"/insitu/summary.csv"))){
    insitu <- read.csv(paste0(gapDir,"/insitu/summary.csv"), header = TRUE)
    summ_table$GRS_in <- round(insitu$GRS, 2)
    summ_table$ERS_in <- round(insitu$ERS, 2)
    summ_table$FCS_in <- round(insitu$FCS, 2)
  }
  
  
  
  if(file.exists(paste0(gapDir,"/combined/fcs_combined.csv"))){
    combined <- read.csv(paste0(gapDir,"/combined/fcs_combined.csv"), header = TRUE)
    summ_table$FCSc_min <- round(combined$FCSc_min, 2)
    summ_table$FCSc_max <- round(combined$FCSc_max, 2)
    summ_table$FCSc_mean <- round(combined$FCSc_mean, 2)
    summ_table$Priority_category <- as.character(combined$FCSc_mean_class)
  }
  
  return(summ_table)
  
  }

 

species_table<-lapply(1:length(speciesList), function(i){
  pt<-species_summary(speciesList[i])
  return(pt)
  
  })

species_table <- do.call(rbind, species_table)
utils::write.csv(species_table, paste0(baseDir, "/species_summary_",Sys.Date() ,".csv"), row.names = FALSE, quote = FALSE)
DT::datatable(species_table, options = list(pageLength = 25))

```



<br>
<br>
<br>
  
  
## Median Model Runs
Evaluation statistics for the median model run for species that enough occurence points to be modeled.

```{r  echo=FALSE, message=FALSE, warning=FALSE}
n= 1
vector <- c()
### need to put is a clause incase
for(i in speciesList){
  csv <- list.files(path = paste0(baseDir,"/",i,"/",run_version,"/"), pattern = '.csv', recursive = TRUE, full.names = TRUE)
  thres <- include(csv, "eval_metrics.csv")
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}
df <- lapply(vector, read.csv)%>%
  bind_rows
df <-df[-c(8),]
write.csv(df, paste0(baseDir, "/median_summary_",Sys.Date() ,".csv"), row.names = FALSE, quote = FALSE)
DT::datatable(df, options = list(pageLength = 25))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
## Gather raw Occurence data for Genus 

n= 1
vector <- c()
### need to put is a clause incase
for(i in speciesList){
  csv <- list.files(path = paste0(baseDir,"/",i,"/",run_version,"/"), pattern = '.csv', recursive = TRUE, full.names = TRUE)
  thres <- include(csv, "rawDataForNA.csv")
  if(length(thres) > 0){
    vector[n] <- thres
    n = n+1
  }
}
df <- lapply(vector, read.csv)%>%
  bind_rows

write.csv(df, paste0(baseDir, "/summaryDocs/occurenceInNA",Sys.Date() ,".csv"), row.names = FALSE, quote = FALSE)
```
<br>
<br>
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE}
if(length(lowOccurence>0)){
  print("theses specices had no acceptable occurance data")
### Species No Run due to low occurence data 
print(lowOccurence )
}

if(length(notModeled>0)){
  print("there species had less then 20 occurence point and could now be modeled")
### Species not randue to Less then 20 occurences. 
print(notModeled)
}

```

```{r  echo=FALSE, message=FALSE, warning=FALSE}
# compile all the modeled Bio Data
n= 1
vector <- c()
  ### need to put is a clause incase
for(i in speciesList){
csv <- list.files(path = paste0(baseDir,"/",i,"/",run_version,"/"), pattern = '.csv', recursive = TRUE, full.names = TRUE)
bioValues <- include(csv, "bioValuesForPresencePoints.csv")
if(length(bioValues) > 0){
vector[n] <- bioValues
n = n+1
}
}

df <- lapply(vector, read.csv)%>%
  bind_rows

write.csv(df, paste0(baseDir, "/modeledBioValues_",Sys.Date() ,".csv"), row.names = FALSE)
```

```{r  echo=FALSE, message=FALSE, warning=FALSE}
# compile all the variable selection options
n= 1
vector <- c()
  ### need to put is a clause incase
for(i in speciesList){
csv <- list.files(path = paste0(baseDir,"/",i,"/",run_version,"/"), pattern = '.csv', recursive = TRUE, full.names = TRUE)
selectVar <- include(csv, "predictorImportance.csv")
if(length(selectVar) > 0){
vector[n] <- selectVar
n = n+1
}
}

df <- lapply(vector, read.csv)%>%
  bind_rows
write.csv(df, paste0(baseDir, "/variablesModeled_",Sys.Date() ,".csv"), row.names = FALSE)
```


